long_ts

### 2019-08-07 SACTN_lts.Rmd
SST.update
Added some code on how to perform a couple of additional tasks.

# Detectig frequency and intensity of upwelling events - Use MHW algorithm 
  Long time series of 30years
  Use OISST 
  Match this up with MCS occuring in the region
  Paper 1 (Local paper)
  Upwelling: 30th percentiles within EBUS

@@

library(tidyverse)
library(plyr)
library(lubridate)
library(ggpubr)
library(zoo)
library(data.table)
library(heatwaveR)
library(viridis)
library(gridExtra)
library(fossil)
library(plyr) # Never load plyr when also loading the tidyverse. It causes a lot of conflicts.
library(stringr)
library(doMC); doMC::registerDoMC(cores = 4)
library(fasttime)
library(xtable)
library(ggpubr)
library(ggpubr); theme_set(theme_pubr())# library(devtools)

```

@@

```{r}
load("~/Documents/Masters_2019/Upwelling_MCS/Data/SACTN_daily_v4.2.RData")
load("~/Documents/Masters_2019/Upwelling_MCS/Data/site_list_v4.2.RData")

@@

# Long time series in_situ temperature
  First creating datasets for each region
  Now I determine the frequency and intensity of MCS using the MHW algorithm

```{r}

filter_SACTN <- function(df){
  SACTN_PN <- SACTN_US %>%
  filter(site == df) %>%
  select(-index, -site, -src, -length) %>% 
  dplyr::rename(t = date) %>% 
  drop_na()
  return(SACTN_PN)
}

SACTN_LB <- filter_SACTN(df = "Lamberts Bay") 
SACTN_PN <- filter_SACTN(df = "Port Nolloth") 
SACTN_SB <- filter_SACTN(df = "Saldanha Bay") 
SACTN_SP <- filter_SACTN(df = "Sea Point") 
SACTN_HB <- filter_SACTN(df = "Hout Bay") 
```

  The rate of onset of MCS can be determined here

```{r}

#' # Or if one wants to calculate MCSs

MCS_func <- function(df){
  res_clim <- ts2clm(df, climatologyPeriod = c("1974-01-01", "2015-12-31"),
                   pctile = 10)
  return(res_clim)
}

PN_MCS <- MCS_func(df = SACTN_PN)
LB_MCS <- MCS_func(df = SACTN_LB)
SB_MCS <- MCS_func(df = SACTN_SB)
SP_MCS <- MCS_func(df = SACTN_SP)

res_clim <- ts2clm(SACTN_HB, climatologyPeriod = c("1974-01-01", "2005-10-19"), # Climatological period differs to the rest
                   pctile = 10)

## Events
MCS_event <- function(df){
  detect <- detect_event(df, coldSpells = TRUE)
  return(detect)
}

PN_MCS_events <- MCS_func(df = PN_MCS)
LB_MCS_events <- MCS_func(df = LB_MCS)
SB_MCS_events <- MCS_func(df = SB_MCS)
SP_MCS_events <- MCS_func(df = SP_MCS)

HB_MCS_events <- detect_event(res_clim, coldSpells = TRUE) # line 118 -
```

@@

```{r}
upwell_func <- function(df){
  res_clim <- ts2clm(df, climatologyPeriod = c("1974-01-01", "2015-12-31"),
                   pctile = 30)
  return(res_clim)
}

PN_upwell <- upwell_func(df = SACTN_PN)
LB_upwell <- upwell_func(df = SACTN_LB)
SB_upwell <- upwell_func(df = SACTN_SB)
SP_upwell <- upwell_func(df = SACTN_SP)

res_clim <- ts2clm(SACTN_HB, climatologyPeriod = c("1974-01-01", "2005-10-19"), # Climatological period differs to the rest
                   pctile = 10)

## Events
upwell_event <- function(df){
  detect <- detect_event(df, minDuration = 2, coldSpells = TRUE)
  return(detect)
}

PN_upwell_events <- MCS_func(df = PN_upwell)
LB_upwell_events <- MCS_func(df = LB_upwell)
SB_upwell_events <- MCS_func(df = SB_upwell)
SP_upwell_events <- MCS_func(df = SP_upwell)

HB_upwell_events <- detect_event(res_clim, minDuration = 2, coldSpells = TRUE) # line 1
```

>>>>

# Detectig frequency and intensity of upwelling events - Use MHW algorithm 
  Long time series of 30 years
  Use OISST 
  Match this up with MCS occuring in the region
  Paper 1 (Local paper)
  Upwelling: 30th percentiles within EBUS

@@

library(tidyverse)
library(lubridate)
library(ggpubr); theme_set(theme_pubr())
library(zoo)
library(data.table)
library(heatwaveR)
library(viridis)
library(gridExtra)
library(fossil)
library(stringr)
library(doMC); doMC::registerDoMC(cores = 4)
library(fasttime)
library(xtable)

# Disable scientific notation for numeric values
# I just find it annoying
options(scipen = 999)
```

@@

```{r}
load("Data/SACTN_daily_v4.2.RData")
load("Data/site_list_v4.2.RData")

@@

# Long _in situ_ temperature time series

Determine the frequency and intensity of MCS using the MHW algorithm

```{r}
# Use purrr::nest() to allow the use of purrr::map() to grouped data
SACTN_MCS <- SACTN_US %>% 
  dplyr::rename(t = date) %>%
  group_by(site) %>% 
  nest() %>% 
  mutate(clims = purrr::map(data, ts2clm,
                          # NB: A 30 year period should be used for a climatology
                          # Conveniently, this period works with all of the time series
                          climatologyPeriod = c("1974-01-01", "2003-12-31"),  pctile = 10),
         events = purrr::map(clims, detect_event, coldSpells = TRUE)) %>% 
  select(-data, -clims)

# Unnest the climatology results
SACTN_MCS_clims <- SACTN_MCS %>% 
  unnest(events) %>%
  filter(row_number() %% 2 == 1) %>%
  unnest(events)

# Unnest the event results
SACTN_MCS_events <- SACTN_MCS %>% 
  unnest(events) %>%
  filter(row_number() %% 2 == 0) %>%
  unnest(events)
```

@@

```{r}
SACTN_upwell <- SACTN_US %>% 
  dplyr::rename(t = date) %>%
  group_by(site) %>% 
  nest() %>% 
  mutate(clims = purrr::map(data, ts2clm,
                          # NB: A 30 year period should be used for a climatology
                          # Conveniently, this period works with all of the time series
                          climatologyPeriod = c("1974-01-01", "2003-12-31"),  pctile = 30),
         events = purrr::map(clims, detect_event, coldSpells = TRUE, minDuration = 2)) %>% 
  select(-data, -clims)

# Unnest the climatology results
SACTN_upwell_clims <- SACTN_upwell %>% 
  unnest(events) %>%
  filter(row_number() %% 2 == 1) %>%
  unnest(events)

# Unnest the event results
SACTN_upwell_events <- SACTN_upwell %>% 
  unnest(events) %>%
  filter(row_number() %% 2 == 0) %>%
  unnest(events)
```

# Test visuals

The figures below compare the largest event detected with the classic MCS definition vs. the upwelling definition used here (i.e. 30th perc. thresh. and 2 day minimum). This isn't very usfeul... but we do see that only 2 of the 5 time series have largest events match up with the different definitions. I'm not sure that is useful though...

```{r}
# Port Nolloth
ggarrange(event_line(SACTN_MCS$events[[1]]),
          event_line(SACTN_upwell$events[[1]]), ncol = 1, nrow = 2)

# Lamberts Bay
ggarrange(event_line(SACTN_MCS$events[[2]]),
          event_line(SACTN_upwell$events[[2]]), ncol = 1, nrow = 2)

# Saldanha Bay
ggarrange(event_line(SACTN_MCS$events[[3]]),
          event_line(SACTN_upwell$events[[3]]), ncol = 1, nrow = 2)

# Sea Point
ggarrange(event_line(SACTN_MCS$events[[4]]),
          event_line(SACTN_upwell$events[[4]]), ncol = 1, nrow = 2)

# Hout Bay
ggarrange(event_line(SACTN_MCS$events[[5]]),
          event_line(SACTN_upwell$events[[5]]), ncol = 1, nrow = 2)
```

The following figures show each of the 5 long time series with the MCS days highlighted in blue, and the upwelling days shown with a rug plot. We see that the 30th percentile threshold shows near constant upwelling, which is likely not accurate. A more selective threshold should probably be used, but this will depned on what the upweliing index results show.

```{r}
ggplot(data = SACTN_MCS_clims, aes(x = t, y = temp)) +
  geom_line() +
  geom_point(data = filter(SACTN_MCS_clims, event == TRUE), colour = "navy") +
  geom_rug(data = filter(SACTN_upwell_clims, event == TRUE), sides = "b") +
  scale_x_date(expand = c(0, 0)) +
  facet_wrap(~site, ncol = 1) +
  theme_grey() +
  labs(y = "Temperature (°C)", x = NULL)
```

# Linear regressions

Below we use `purrr::map()` to apply a linear regression to every variable from our MCS and upwelling results. These results may need some thinking. I'm not certain that the negative values inherent in the MCS results are being handled correctly here.

```{r}
# Convenience function for below
lm_coeff <- function(df){
  res <- lm(formula = val ~ date_peak, data = df)
  res_coeff <- round(as.numeric(res$coefficient[2]), 4)
}

# Changes in MCS metrics
SACTN_MCS_lm <- SACTN_MCS_events %>% 
  select(-c(index_start:index_end)) %>% 
  gather(key = "var", value = "val", -c(site:date_end)) %>% 
  group_by(site, var) %>% 
  nest() %>% 
  mutate(slope = purrr::map(data, lm_coeff)) %>% 
  select(-data) %>% 
  unnest() %>% 
  # convert from daily to decadal values
  mutate(slope = slope * 365.25*10)

# Changes in upwelling metrics
SACTN_upwell_lm <- SACTN_upwell_events %>% 
  select(-c(index_start:index_end)) %>% 
  gather(key = "var", value = "val", -c(site:date_end)) %>% 
  group_by(site, var) %>% 
  nest() %>% 
  mutate(slope = purrr::map(data, lm_coeff)) %>% 
  select(-data) %>% 
  unnest() %>% 
  # convert from daily to decadal values
  mutate(slope = slope * 365.25*10)
```
